Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-02 method.tex
Removed some more continuous time notation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@849 ccdd05d5-56c1-461a-bf98-85f6876523b4

The monotonic time-frequency relationship of \eqref{eq:fgw} allows us to choose
time-slice boundaries that require substantially less bandwidth at early times
in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
of the waveform by computing part of the filter integral at a lower sample
rate.  Specifically we consider here time slice boundaries with the next
highest power-of-two sample rate that critically samples the time sliced
template.  The time slices consist of the $S$ intervals $\left(t^S,
t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$ sampled at frequencies
$f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time sliced templates may be downsampled without aliasing, so we define them as
%
\begin{equation}
\label{eq:time-sliced-templates}
h_{i}^{s}[k] \equiv
	\cases{
		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
		0 & otherwise.
	}
\end{equation}
%
An example time-slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
%
\begin{table}[h!]
\begin{minipage}[c]{0.52\textwidth}
\centering
\vspace{0.8cm}
\includegraphics{time_slices.pdf}
\end{minipage}
\begin{minipage}[c]{0.3\textwidth}
\centering
\input{time_slices.tex}
\end{minipage}
\caption{\label{table:time_slices} Example of nearly critically sampled,
power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
with a time frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
preceeding coalescence and \slicessamps\ are the number of sample points in the
$s^{\mathrm{th}}$ filter.}
\end{table}
%

\editorial{Need to convert some continuous time notation to discrete time here.}Rather than applying a unique time-slice decomposition to each template
waveform, we find a decomposition that is adequate for the entire set
$\{x_i\}$.  This is facilated by choosing templates in the set $\{x_i\}$ with
similar chirp masses, $\mathcal{M}$.  The time-slice decompositions of the
filters $x_i(\tau)$ lead to new filters $y_{ik}(\tau)$ satisfying
%
\begin{equation}
\label{eq:time_slices}
x_i(\tau) \approx \sum_k^{\numslices} y_{ik}(\tau)
\end{equation}
where \numslices\ is the number of time slices required.  We note that
the time slice decomposition in equation~(\ref{eq:time-slices}) is manifestly
orthogonal since the time slices are disjoint in time.  In the next
section we examine how to reduce the number of filters within each time slice via
singular value decomposition of the time sliced templates.

>>>>

The monotonic time-frequency relationship of \eqref{eq:fgw} allows us to choose
time-slice boundaries that require substantially less bandwidth at early times
in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
of the waveform by computing part of the filter integral at a lower sample
rate.  Specifically we consider here time slice boundaries with the next
highest power-of-two sample rate that critically samples the time sliced
template.  The time slices consist of the $S$ intervals $\left(t^S,
t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$ sampled at frequencies
$f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time sliced templates may be downsampled without aliasing, so we define them as
%
\begin{equation}
\label{eq:time-sliced-templates}
h_{i}^{s}[k] \equiv
	\cases{
		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
		0 & otherwise.
	}
\end{equation}
%
An example time-slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
%
\begin{table}[h!]
\begin{minipage}[c]{0.52\textwidth}
\centering
\vspace{0.8cm}
\includegraphics{time_slices.pdf}
\end{minipage}
\begin{minipage}[c]{0.3\textwidth}
\centering
\input{time_slices.tex}
\end{minipage}
\caption{\label{table:time_slices} Example of nearly critically sampled,
power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
with a time frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
preceeding coalescence and \slicessamps\ are the number of sample points in the
$s^{\mathrm{th}}$ filter.}
\end{table}

Since waveforms with neighboring intrinsic source parameters $\vec\theta$ have similar
time-frequency evolution, it is possible to design computationally efficient time slices
for an extended region of parameter space rather than to design different time slices for
each template.

We note that
the time slice decomposition in equation~(\ref{eq:time-slices}) is manifestly
orthogonal since the time slices are disjoint in time.  In the next
section we examine how to reduce the number of filters within each time slice via
singular value decomposition of the time sliced templates.
